---
title: "K-means clustering - step by step"
output: 
  html_document: 
    toc: yes
    toc_float:
      toc_collapsed: yes
    number_sections: yes
    theme: paper
    df_print: kable
---


# Preliminary steps

Load required packages:

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(FactoMineR)
library(readxl)
library(readr)
```

and data (rows are individual, columns are variables):

```{r load-data, message=FALSE, warning=FALSE}
# # df <- USArrests[,1:2]
# df_birds <- read_xlsx("data/short.xlsx", col_types = c("text","numeric","numeric"))
# df <- na.omit(df_birds)
# df <- scale(df[,2:3])
# head(df)

european_birds <- read_delim("data/european_birds.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
european_birds %>% 
  dplyr::select(Species,
                LengthU_MEAN,
                # WingU_MEAN,
                Clutch_MEAN,
                `Life span`) -> df_birds
df_birds <- na.omit(df_birds)
df <- scale(df_birds[,2:dim(df_birds)[2]])
head(df)

```


Visualize the dissimilarities:

```{r dissimilarities-plot}
distance <- get_dist(df)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```


# Define the number of clusters

with the Elbow method

```{r cl-def-elbow}
fviz_nbclust(df, kmeans, method = "wss")
```

or the Average silhouette method

```{r cl-def-sil}
fviz_nbclust(df, kmeans, method = "silhouette")
```


or the Gap statistics

```{r cl-def-gap}
gap_stat <- clusGap(df, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
```


# Compute k-means clusering

using the chosen number of clusters

```{r clustering}
# Compute k-means clustering with k = 4
set.seed(123)
final <- kmeans(df, 3, nstart = 25)
print(final)
```

```{r cl-viz}
fviz_cluster(final, data = df,
             geom = c("point"),
             ellipse.type = "euclid")
```


```{r cl-list, include=FALSE}
final$cluster
```

Each species is associated to a cluster, and the results are saved to csv and xlsx files:

```{r results}
df_birds %>% 
  mutate(cluster_id = final$cluster) -> df_birds_cluster
write.csv(df_birds_cluster, file = "output/df_birds_cluster.csv")
writexl::write_xlsx(df_birds_cluster, "output/df_birds_cluster.xlsx")
head(df_birds_cluster)
```

This is a summary of the average features for each cluster:

```{r}
df_birds_cluster %>% 
  group_by(cluster_id) %>% 
  summarise(meanLength = mean(LengthU_MEAN),
            # meanWing = mean(WingU_MEAN),
            meanLifeSpan = mean(`Life span`),
            meanClutch = mean(Clutch_MEAN))
```



